---
title: "An overview of plotly's R API"
author: "Carson Sievert"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteIndexEntry{Plotly DSL}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  fig.width = 7, 
  fig.height = 3
)
```

To create a plotly visualization, start with `plot_ly()`.

```{r}
library(plotly)
plot_ly(economics, x = date, y = unemploy / pop)
```

A plotly visualization is composed of one (or more) trace(s), and every trace has a `type` (the default type is 'scatter'). The arguments/properties that a trace will respect ([documented here](https://plot.ly/r/reference)) depend on it's type. A scatter trace respects `mode`, which can be any combination of "lines", "markers", "text" joined with a "+":

```{r}
library(plotly)
plot_ly(economics, x = date, y = unemploy / pop, 
        type = "scatter", mode = "markers+lines")
```

You can manually add a trace to an existing plot with `add_trace()`. In that case, you'll want to either `name` your traces, or hide the legend by setting `showlegend = FALSE`

```{r}
m <- loess(unemploy / pop ~ as.numeric(date), data = economics)
p <- plot_ly(economics, x = date, y = unemploy / pop, name = "raw") 
add_trace(p, y = fitted(m), name = "loess")
```

__plotly__ was designed with a [pure, predictable, and pipeable interface](https://dl.dropboxusercontent.com/u/41902/pipe-dsls.pdf) in mind, so you can also use the `%>%` operator to create a visualization pipeline:

```{r}
economics %>%
  plot_ly(x = date, y = unemploy / pop) %>% 
  add_trace(y = fitted(m)) %>%
  layout(showlegend = F)
```

Furthermore, `plot_ly()`, `add_trace()`, and `layout()`, all accept a data frame as their first argument and output a data frame. As a result, we can inter-weave data manipulations and visual mappings in a single pipeline.

```{r}
economics %>%
  transform(rate = unemploy / pop) %>%
  plot_ly(x = date, y = rate) %>% 
  subset(rate == max(rate)) %>%
  layout(
    showlegend = F, 
    annotations = list(x = date, y = rate, text = "Peak", showarrow = T)
  )
```

Although these functions output special "plotly data frames", `plot_ly()` doesn't require a data frame for input. This makes chart types that accept a `z` argument especially easy to use if you have a numeric matrix:

```{r, eval = FALSE}
s <- plot_ly(z = volcano, type = "surface")
```

If you have [a free plotly account](https://plot.ly/r/getting-started/), `plotly_POST()` makes it easy to host/share any plotly figure.

```{r, eval = FALSE}
plotly_POST(s)
```

```{r, results = 'asis', echo = FALSE}
cat(plotly:::plotly_iframe("https://plot.ly/~agvd/1794/", 600, 600))
```


## Special arguments

The __plotly__ function arguments used thus far are documented in [the figure reference](https://plot.ly/r/reference), but there are a handful of additional arguments not described in the reference that make common visualizations a bit easier. These arguments are very much inspired by the semantics of ggplot2's `qplot()` in the sense that a scales are automatically applied these variables.

### The color argument

#### Qualitative color mappings

If a ordinal variable (aka a non-ordered factor variable) is assigned to color, then a qualitative color palette is used by default.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = Species, mode = "markers")
```

If you want to change the default palette, it's recommended that you provide a <http://colorbrewer2.org> qualitative pallette name (e.g., "Set1" or "Accent") to the colors argument.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = Species, colors = "Set1", mode = "markers")
```

In this case, the palette consists of 9 colors and the default behavior is to pick colors that are furthest apart ("#E41A1C", "#FF7F00", and "#999999").

```{r}
cols <- RColorBrewer::brewer.pal(9, "Set1")
scales::show_col(cols)
```

If you'd like more control over the mapping, you can provide a vector of colors (of appropriate length).

```{r}
cols <- RColorBrewer::brewer.pal(nlevels(iris$Species), "Set1")
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = Species, colors = cols, mode = "markers")
```

#### Sequential color mappings

If either a numeric or an ordered factor is mapped to color, `plot_ly()` applies a sequential color scale by default.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = as.ordered(Species), mode = "markers")
```

In the case of continuous numeric variables, `plot_ly()` performs a linear mapping between the data and an interpolated color pallette.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = Sepal.Length, mode = "markers")
```

The colors argument takes arbitrary color codes of arbitrary length. Here is how we could use it to replicate the default mapping in ggplot2.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = Sepal.Length, colors = c("#132B43", "#56B1F7"), 
        mode = "markers")
```

#### Diverging color mappings

To obtain a diverging color mapping, just provide a diverging palette to the colors argument.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        color = Sepal.Length, colors = "PuOr", mode = "markers")
```

### The symbol argument

To encode values using symbols, use the symbol argument.

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        symbol = Species, mode = "markers")
```

To change the default symbols used, use the symbols argument. All the valid symbol types are listed [here](https://plot.ly/r/reference/#marker).

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, mode = "markers",
        symbol = Species, symbols = c("cross", "square", "triangle-down"))
```


### The group argument and `subplot()`

Using the group argument splits the data into different plotly "traces".

```{r}
plot_ly(iris, x = Petal.Length, y = Petal.Width, 
        group = Species, mode = "markers")
```

Although we haven't specified a coloring scheme, plotly will employ one on it's own default scheme. The group argument is quite powerful when used in conjunction with `subplot()` in order to anchor traces onto different axes.

```{r}
iris$id <- as.integer(iris$Species)
p <- plot_ly(iris, x = Petal.Length, y = Petal.Width, group = Species,
             xaxis = paste0("x", id), mode = "markers")
subplot(p)
```

Since `subplot()` does not assume x/y axes are on a common scale, it does not impose any restrictions on the range by default. However, you can change this by pre-specifying the range of the [axis objects](https://plot.ly/r/reference/#xaxis) via the `layout()` function.

```{r}
p2 <- layout(
  p, 
  xaxis = list(range = range(Petal.Length) + c(-0.1, 0.1)),
  yaxis = list(range = range(Petal.Width) + c(-0.1, 0.1))
)
subplot(p2)
```

Part of the magic of `subplot()` is that it generates axis objects with appropriate anchor and domain properties. After generating a subplot, you can always reference these axis objects to customize each plot.

```{r}
layout(
    subplot(p2),
    yaxis2 = list(title = ""), 
    yaxis3 = list(title = "")
)
```


[See here](https://plot.ly/r/map-subplots-and-small-multiples/) for another example of using the group argument to make small multiples (with maps!).
