---
title: "The `subplot()` function"
author: "Carson Sievert"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteIndexEntry{subplot}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  fig.width = 4, 
  fig.height = 2
)
```

The `subplot()` function provides a flexible interface for arranging multiple plots in a single view. There are a few different ways to use `subplot()`, but the simplest way to pass plotly visualization objects directly to `subplot()`.

```{r}
library(plotly)
p1 <- plot_ly(economics, x = date, y = unemploy)
p2 <- plot_ly(economics, x = date, y = uempmed)
subplot(p1, p2)
```

This particular subplot could be improved if we share the x-axis so that zooming and panning events are synchronized. It also makes sense to keep axis titles and remove the redundant legend (since `subplot()` returns a plotly visualization object, any [layout attribute](https://plot.ly/r/reference/#layout) can be altered via `layout()`).

```{r}
s <- subplot(p1, p2, nrows = 2, shareX = TRUE, titleY = TRUE)
layout(s, showlegend = FALSE)
```

By default, every subplot is allocated equal spacing, and the margins between them are set somewhat arbitrarily, but that can easily be altered:

```{r}
map <- plot_ly(
  z = state.area, text = state.name, locations = state.abb,
  type = 'choropleth', locationmode = 'USA-states', geo = "geo"
)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  lakecolor = toRGB('white')
)
map <- layout(map, geo = g)
subplot(
  plot_ly(x = state.name, y = state.area, type = "bar"), map,
  nrows = 2, margin = c(0, 0, 0.25, 0), heights = c(0.2, 0.8)
)
```

In addition to a `heights` argument for controlling the proportional height of each row, there is also a `widths` argument for controlling the proportional width of each column:

```{r}
x <- rnorm(100)
y <- rnorm(100)
m <- list(color = "black")
subplot(
  plot_ly(x = x, type = "histogram", marker = m), 
  plotly_empty(), 
  plot_ly(x = x, y = y, mode = "markers",  marker = m), 
  plot_ly(y = y, type = "histogram", marker = m),
  nrows = 2, heights = c(0.2, 0.8), widths = c(0.8, 0.2), 
  shareX = TRUE, shareY = TRUE
) %>% layout(showlegend = FALSE)
```

The `subplot()` function also understands ggplot2 objects and converts them via `ggplotly()`.

```{r, fig.height = 8}
e <- tidyr::gather(economics, variable, value, -date)
gg1 <- ggplot(e, aes(date, value)) + geom_line() +
  facet_wrap(~variable, scales = "free_y", ncol = 1)
gg2 <- ggplot(e, aes(factor(1), value)) + geom_violin() +
  facet_wrap(~variable, scales = "free_y", ncol = 1) + 
  theme(axis.text = element_blank(), axis.ticks = element_blank())
subplot(gg1, gg2) %>% layout(margin = list(l = 50))
```




