---
title: "Render plotly MathJax inside rmarkdown"
output: html_document
---

Some HTML-based MathJax:

$$ \alpha+\beta $$

You _could_ print this this **plotly** graph with SVG-based rendering, but it would break the HTML-based rendering of **rmarkdown**!

```{r message = FALSE}
library(plotly)

p <- plotly_empty() %>%
  add_trace(x = 1, y = 1, text = "$\\alpha$", mode = "text", size = I(1000)) %>%
  config(mathjax = "cdn")
```

Instead, use something like the **widgetframe** package to create a responsive iframe. This ensures the SVG-based rendering that plotly requires is done in isolation (i.e., independent of **rmarkdown**'s HTML-based rendering).

```{r}
widgetframe::frameableWidget(p)
```

Or, do it the old-fashioned way by saving your plotly graph to an HTML file and use an HTML `<iframe>` to 

```{r}
htmlwidgets::saveWidget(p, "my-plotly-plot.html")
```


<iframe src="my-plotly-plot.html" width="100%" height="400" id="igraph" scrolling="no" seamless="seamless" frameBorder="0"> </iframe>
